---
title: "What next?"
subtitle: "Probably something fundamental like linear regression."
author: Ángela Castillo-Gill
date: '2018-11-29'
slug: what-next
categories:
  - Journal
tags: 
  - Learning
draft: FALSE
summary: "Probably something fundamental like linear regression."
output:
 blogdown::html_page:
  fig_caption: true
editor_options: 
  chunk_output_type: console
---

# Daily goals:

1. Reflect on what post I would like do next, what tool will I use, what dataset will I use and why?

```{r read data, echo=FALSE, message=FALSE}
library(tidyverse)
library(rpart)
library(rpart.plot)
library(knitr)
options(digits=4)
options(scipen = 999)
opts_chunk$set(echo = FALSE,
               warning = FALSE,
               error = FALSE,
               message = FALSE,
               collapse= TRUE,
               comment = NA,
               tidy = TRUE)
theme_set(theme_light())
homes <- readRDS(file="/Volumes/TOSHIBAEXT/google_drive/acastillogill/content/homes.rds")
```

The last posts I haven't planned properly in terms of what tool I'm going to use, why is it well-suited, and what dataset would I like to use. Don't get me wrong, there are some positive aspects to this lacking of planning. I get to see lots of raw datasets where I can see the challenges of dealing with raw data. 

Despite linear regression being such a foundational thing, I haven't used it in any post so far. I did attempt to do it with the Kiva loans one, but the dataset was so large that it got a little bit out of hand. Performing linear regression on 2,000 observations is very much different than doing so on a dataset with 500,000. I rememeber the distributions being bimodal and that tripping me up and also the number of outliers at a time. 

I asked for help at some point and someone on CrossValidated, I think quite appropriately so, said: stop and take baby steps. And I didn't want to but now I kind of see why. The frustration of taking very large steps can exhaust and feel daunting. It's much better to take small steps and come across issues slowly that seem figure-out-able. 

So for this next post I would like to use a linear regression. Why? Because I would like to properly test my understanding of the assumptions it carries, what are the steps I should take, how will I deal with outliers, and how I will I choose the best model and why.

Getting linear regression is critical/crucial/essential in this whole data science thing and I intend to get it right. 

Tuesday and Wednesday were this month's R-Ladies reading group and it was bloody EXCELLENT. I learned so much about the importance of being organised when doing computing and how to do so. I'm having a ball reading Hadley Wickam's reading list and learning lots. I've also started the book Introduction to Statistical Learning which is the foundation of these blogpost.

Ok, moving on. 

I want to apply linear regression because it is critical I get right, but what will be the dataset and how do I come up with a question? Ah. Yes. A bit more tricky. The reason it's a bit more tricky is because receiving a simple dataset and cleaning it, getting to know it, etc, is just a pre step to figuring out what is the question we want to ask.

Oh, just had an idea, why don't I just use the same one? I could use the property price and get on with it. It could help look at the effects of adjusting for another variable and also, how to deal with dummy variables. Hmmmm.

Could be a good idea, since it's a dataset I know well, I'm sort of tempted. Also, it could be a good start to compare the results with property values. I definitely need to improve my interpretation of results. It's a dataset I understand well so why not. Ok. Same dataset. I wanted to go through the Tidy Tuesday one but maybe in another post. I can't move on until I get linear regression right. I mean I can but I don't want to.

So, what is the first condition for a linear regression:

-The input ($X_i$) and the output ($Y_i$) do not have to be normally distributed. 
-The $Y|X$ do have to be linearly related. 

```{r}

```

