---
title: "Describing outliters"
subtitle: "The US, China, India, big impact."
author: Ángela Castillo-Gill
date: '2018-12-05'
slug: adding-variables
categories:
  - Journal
tags: 
  - Learning
draft: FALSE
summary: "The US, China, India, big impact."
output:
 blogdown::html_page:
  fig_caption: true
editor_options: 
  chunk_output_type: console
---

```{r setup, echo=FALSE, message=FALSE}
library(tidyverse)
library(dplyr)
library(knitr)
library(here)
library(scales)
library(magrittr)
options(digits=4)
options(scipen = 999)
opts_chunk$set(echo = TRUE,
               warning = FALSE,
               error = FALSE,
               message = FALSE,
               collapse= TRUE,
               comment = NA,
               tidy = TRUE)
theme_set(theme_light())
r_downloads <- readRDS(file=here::here("static","data","r-downloads.rds"))
```



[Data source](https://github.com/rfordatascience/tidytuesday/tree/master/data/2018-10-30)

```{r downloads-by-country}
r_downloads%>%
  group_by(country)%>%
  summarise(n=n())%>%
  arrange(desc(n))->downloads_per_country
```

Now I'm going to give the countries full names, to do this I will use the ISO code to merge and obtain names. 

```{r read-country-codes}
iso_codes <- read_csv(here("static","data","iso_countries.csv"))
```

```{r merge-country-codes}
downloads_per_country%<>%
  right_join(iso_codes, by=c("country"="iso"))
```


## Which are the top ten countries in terms of downloads
```{r top-20-countries-plot}
downloads_per_country%>%
  arrange(desc(n))%>%
  filter(!is.na(name))%>%
  head(20)%>%
  mutate(name=fct_reorder(name,n))%>%
  ggplot(aes(name,n,fill=name))+
  geom_col()+
  expand_limits(y=0)+
  coord_flip()+
  scale_y_continuous(labels=comma_format())+
  labs(title="Top 20 countries by R downloads",
       x="Country",
       y="Downloads")+
  theme(legend.position = "none")
  
```
Today I've added one variable `internet_users` which sums the total internet users per country. The point is to slowly add regressors or independent variables and see if they have a linear relationship with `r-downloads`. If yes, I'll add them to a model.

```{r read-internet-users}
internet_users <-read_csv(here("static","data","internet_users.csv"))
```

```{r merge-internet-users}
downloads_users_country <-downloads_per_country%>%
  right_join(internet_users, by=c("name"="name"))%>%
  filter(!is.na(name),
         !is.na(users),
         !is.na(n))%>%
  arrange(desc(n))
  
```

```{r downloads-internet-users-scatter-plot, fig.cap="There are clearly some influential points, US, China and India, that are outliers in case of a regression."}
library(plotly)
library(ggrepel)
downloads_internet_plot<- downloads_users_country%>%
  arrange(desc(n))%>%
  ggplot(aes(users, n, label=name, color=region))+
  #Add light opacity
  geom_jitter(alpha=0.3)+
  #Add nice commas x axis
  scale_x_continuous(labels=comma_format())+
  #Add nice commas y axis
scale_y_continuous(labels=comma_format())+
  #Add a linear regression line for all regions
  geom_smooth(aes(group=1),method="lm")+
  #Fix legend
theme(legend.position = "bottom", legend.title = element_blank())+
   #Add title
  labs(title="Do R-Language downloads increase with more internet users?",
       y="Downloads per country",
       x="Internet users in each country")
  

ggplotly(downloads_internet_plot)%>%
   layout(legend = list(orientation = "h", x = 0.4, y = -0.2))
```

Today I'm going to describe the big outliers in the set and see what their effect is on the regression.

```{r fitting-model-first}
internet_downloads_fit <- lm(n~users, data = downloads_users_country)
summary(internet_downloads_fit)$coef
par(mfrow=c(2,2))
plot(internet_downloads_fit)
```

```{r dataset-without-outliers}
downloads_users_country%>%
  filter(name!="United Kingdom",
         name!="Germany",
         name!="China",
         name!="India",
         name!="United States of America",
         name!="Brazil",
         name!="Japan",
         name!="Canada",
         name!="Australia",
         name!="Namibia")->outliers_gone
```


```{r plot-without-outliers}
downloads_internet_plot_out<- outliers_gone%>%
  arrange(desc(n))%>%
  ggplot(aes(users, n, label=name, color=region))+
  #Add light opacity
  geom_jitter(alpha=0.3)+
  #Add nice commas x axis
  scale_x_continuous(labels=comma_format())+
  #Add nice commas y axis
scale_y_continuous(labels=comma_format())+
  #Add a linear regression line for all regions
  geom_smooth(aes(group=1),method="lm")+
  #Fix legend
theme(legend.position = "bottom", legend.title = element_blank())+
   #Add title
  labs(title="Do R-Language downloads increase with more internet users?",
       subtitle = "China, India, United Kingdom, and Germany removed.",
       y="Downloads per country",
       x="Internet users in each country")
  

ggplotly(downloads_internet_plot_out)%>%
   layout(legend = list(orientation = "h", x = 0.4, y = -0.2))
```

```{r fitting-model-without-outliers}
internet_downloads_fit_2 <- lm(n~users, data = outliers_gone)
summary(internet_downloads_fit_2)$coef
par(mfrow=c(2,2))
plot(internet_downloads_fit_2)
```


