---
title: "Interactive maps with Tmap!"
subtitle: "Downloads per capita and total downloads, up!"
author: Ángela Castillo-Gill
date: "`r format(Sys.Date())`"
slug: interactive-maps-with-tmap
categories:
  - Journal
tags: 
  - Learning
draft: FALSE
summary: "Downloads per capita and total downloads, up!"
output: 
 blogdown::html_page:
  fig_caption: true
editor_options: 
  chunk_output_type: console
---

```{r setup, echo=FALSE, message=FALSE}
library(dplyr)
library(knitr)
library(here)
library(scales)
library(magrittr)
library(ggplot2)
library(cowplot)
library(tidyverse)
library(Metrics)
library(rsample)
library(broom)
library(sf)
library(rgdal)
library(summarytools)
library(leaflet)
library(tmap)
library(tmaptools)
opts_chunk$set(echo = TRUE,
               warning = FALSE,
               error = FALSE,
               message = FALSE,
               collapse= TRUE,
               comment = NA,
               tidy = TRUE)
theme_set(theme_light())

options(
  digits=3,
  scipen = 999
)


scale_colour_discrete <- function(...) scale_colour_brewer(..., palette="Paired")
scale_fill_discrete <- function(...) scale_fill_brewer(... , palette="Paired")

r_downloads <- readRDS(file=here::here("static","data","r-downloads.rds"))

r_downloads <- na.omit(r_downloads)
```

[Data source](https://github.com/rfordatascience/tidytuesday/tree/master/data/2018-10-30)

```{r downloads-by-country}
r_downloads%>%
  group_by(country)%>%
  summarise(n=n())%>%
  filter(!is.na(n))%>%
  arrange(desc(n))->downloads_per_country
```

Now I'm going to give the countries full names, to do this I will use the ISO code to merge and obtain names. 

```{r read-country-codes}
iso_codes <- read_csv(here::here("static","data","iso_countries.csv"))
```

```{r merge-country-codes}
downloads_per_country%<>%
  full_join(iso_codes, by=c("country"="iso"))
```



## Which are the top ten countries in terms of downloads
```{r plot-top-10-countries}
downloads_per_country%>%
  arrange(desc(n))%>%
  head(10)%>%
  mutate(name=fct_reorder(name,n))%>%
  ggplot(aes(name,n,fill=name))+
  geom_col()+
  expand_limits(y=0)+
  coord_flip()+
  scale_y_continuous(labels=comma_format())+
  labs(title="Top 10 countries by R downloads",
       x="",
       y="Downloads")+
  theme(legend.position = "none")
  
```
Today I've added one variable `internet_users` which sums the total internet users per country. The point is to slowly add regressors or independent variables and see if they have a linear relationship with `r-downloads`. If yes, I'll add them to a model.

```{r read-internet-users}
internet_users <-read_csv(here::here("static","data","internet_users.csv"))
```

```{r merge-internet-users}
downloads_users_country <-downloads_per_country%>%
  right_join(internet_users, by=c("name"="name"))%>%
  filter(!is.na(name),
         !is.na(users),
         !is.na(n))%>%
  arrange(desc(n))
  
```

```{r plot-downloads-internet-users, fig.cap="There are clearly some influential points, US, China and India, that are outliers in case of a regression."}
library(plotly)
downloads_internet_plot<- downloads_users_country%>%
  arrange(desc(n))%>%
  ggplot(aes(users, n, label=name, color=region))+
  #Add light opacity
  geom_point(alpha=0.5)+
  #Add nice commas x axis
  scale_x_continuous(labels=comma_format())+
  #Add nice commas y axis
scale_y_continuous(labels=comma_format())+
  #Add a linear regression line for all regions
  geom_smooth(aes(group=1),method="lm")+
  #Fix legend
theme(legend.position = "bottom", legend.title = element_blank())+
   #Add title
  labs(title="Do R-Language downloads increase with more internet users?",
       y="Downloads per country",
       x="Internet users in each country")
  

ggplotly(downloads_internet_plot)%>%
   layout(legend = list(orientation = "h", x = 0.4, y = -0.2))
```



```{r dataset-without-outliers}
downloads_users_country%>%
  filter(name!="United Kingdom",
         name!="Germany",
         name!="China",
         name!="India",
         name!="United States of America",
         name!="Brazil",
         name!="Japan",
         name!="Canada",
         name!="Australia",
         name!="Namibia")->outliers_gone
```


```{r plot-without-outliers}
downloads_internet_plot_out<- outliers_gone%>%
  arrange(desc(n))%>%
  ggplot(aes(users, n, label=name, color=region))+
  #Add light opacity
  geom_point(alpha=0.5)+
  #Add nice commas x axis
  scale_x_continuous(labels=comma_format())+
  #Add nice commas y axis
scale_y_continuous(labels=comma_format())+
  #Add a linear regression line for all regions
  geom_smooth(aes(group=1),method="lm")+
  #Fix legend
theme(legend.position = "bottom", legend.title = element_blank())+
   #Add title
  labs(title="Do R-Language downloads increase with more internet users?",
       subtitle = "China, India, United Kingdom, and Germany removed.",
       y="Downloads per country",
       x="Internet users in each country")
  

ggplotly(downloads_internet_plot_out)%>%
   layout(legend = list(orientation = "h", x = 0.4, y = -0.2))
```


As I was removing outliers yesterday, I was thinking, what do all these countries I'm removing have in common and what variable can I add to capture some of that variability.

I mean, look at them, United Kingdom, Germany, China, India, USA, Brazil, Japan, Canada, Australia, Namibia... They are in almost all cases "developed" countries. Let's add the Human Development Indicator as a variable to see if it can capture some of that variability. 

```{r read-hdi-index}
human_dev_ind <-read_csv(here::here("static","data","HDI.csv"))
```


```{r merge-hdi-users}
downloads_hdi_country <-downloads_users_country%>%
  right_join(human_dev_ind, by=c("name"="name"))%>%
  filter(!is.na(name),
         !is.na(users),
         !is.na(n),
         !is.na(HDI))%>%
  arrange(desc(HDI))
```

```{r hdi-downloads-plot}
downloads_hdi_plot<- downloads_hdi_country%>%
  ggplot(aes(x=HDI, y=n, label=name, color=region))+
  #Add light opacity
  geom_point(alpha=0.5)+
  #Add nice commas x axis
  scale_x_continuous(labels=percent_format())+
  #Add nice commas y axis
scale_y_continuous(labels=comma_format())+
  #Add a linear regression line for all regions
  geom_smooth(aes(group=1),method="lm")+
  #Fix legend
theme(legend.position = "bottom", legend.title = element_blank())+
   #Add title
  labs(title="Human Development Index (HDI) and downloads.",
       y="Downloads per country",
       x="Human Development Index")

ggplotly(downloads_hdi_plot)
```

Wow! There does seem to be a strong linear relationship between the variables. Now, it might be that countries with more internet users also have a higher human development index. Makes sense? Right? Let me check the correlation:

```{r checking-correlation}
cor(downloads_hdi_country$users,downloads_hdi_country$HDI)*100
```

Hmmmm only 8% I thought it would be higher. Multicollinearity is a threat in regression. Why add another variable that whose explanatory variable is contained in another.

Let me see the model with this new variable in:

```{r fitting-hdi-users}
internet_fit <- lm(n ~ users, data=downloads_hdi_country)
hdi_fit <- lm(n ~ users + HDI, data=downloads_hdi_country)
summary(hdi_fit)$coef
```

So the HDI is a significant variable if our Type error rate is 0.05, but barely. Internet users still is top.
```{r diagnostics-hdi-users}
par(mfrow=c(2,2))
plot(hdi_fit)
```

Wow! They bloody residuals look not that bad. For like the first time ever in my experience! There is more or less equal variation around zero, they do look more or less distributed normally, and of course we have some outliers, we always do. I'm not too inclined to add another variable. Let's look at the adjusted $R^2$.

```{r adj-r-squared}
summary(hdi_fit)$adj.r.squared*100
```

Mmmm, those two variables, internet users and the HDI only explain about 20.53% of the variation. I want to increase the explanatory power of the model, what to do?

So today I would like break up the HDI index into its component parts and see how each separate feature relates to R downloads. 

The HDI is a mix of three elements: 

- Health: Life expectancy (years)
- Education: Expected years of schooling(years)/Mean years of schooling (years)
- Standard of living: Gross national income per capita (2011 PPP $)

So I have downloaded those variables and I will now plot them against downloads and also fit the regression model again. 

```{r read-four-new-indicators}
standard_liv <- read_csv(here::here("static","data","GNIPC.csv"))
standard_liv$GNI <- as.integer(standard_liv$GNI)

life_exp <- read_csv(here::here("static","data","life_exp.csv"))
edu_year <- read_csv(here::here("static","data","edu_year.csv"))
pop <-read_csv(here::here("static","data","country_populations.csv"))
edu_year <- edu_year%>%
  filter(Indicator=="Mean years of schooling (ISCED 1 or higher), population 25+ years, both sexes",
         year==2016)%>%
  select(country,educ)
```


```{r join-new-indicators}
full_hdi <- downloads_hdi_country%>%
  left_join(pop, by=c("name"="name"))%>%
  left_join(standard_liv, by=c("iso_3"="iso"))%>%
  left_join(life_exp, by=c("country_code"="country_code"))%>%
  left_join(edu_year, by=c("iso_3"="country"))%>%
  arrange(name)
```

# Which are the biggest R users per capita?

```{r plot-users-per-capita, fig.cap="Namibia is the country with most R downloads per capita! I'm surprised!"}
full_hdi%>%
  mutate(downloads_per_capita=n/pop_2018)%>%
  arrange(desc(downloads_per_capita))%>%
  head(10)%>%
  mutate(name=fct_reorder(name,downloads_per_capita))%>%
  ggplot(aes(name,downloads_per_capita,fill=name))+
  geom_col()+
  expand_limits(y=0)+
  coord_flip()+
  scale_y_continuous(labels=comma_format())+
  labs(title="Top 10 countries by R downloads per capita",
       x="",
       y="Downloads per capita")+
  theme(legend.position = "none")
```


```{r plot-life-exp}
life_exp_plot<- full_hdi%>%
  ggplot(aes(x=life_exp, y=n, label=name, color=region))+
  #Add light opacity
  geom_point(alpha=0.5)+
  #Add nice commas y axis
scale_y_continuous(labels=comma_format())+
  #Add a linear regression line for all regions
  geom_smooth(aes(group=1),method="lm")+
  #Remove legend
  theme(legend.position = "none")+
   #Add title
  labs(title="Life expectancy at birth and R downloads.",
       y="Downloads per country",
       x="Life expectancy at birth (Years)")

```

```{r educ-plot}
educ_plot<- full_hdi%>%
  ggplot(aes(x=educ, y=n, label=name, color=region))+
  #Add light opacity
  geom_point(alpha=0.5)+
  #Add nice commas y axis
scale_y_continuous(labels=comma_format())+
  #Add a linear regression line for all regions
  geom_smooth(aes(group=1),method="lm")+
  #Fix legend
theme(legend.position = "bottom", legend.title = element_blank())+
   #Add title
  labs(title="Mean years of schooling (years) and R downloads.",
       y="Downloads per country",
       x="Mean years of schooling (years)")

```


```{r GNI_plot}
GNI_plot<- full_hdi%>%
  ggplot(aes(x=GNI, y=n, label=name, color=region))+
  #Add light opacity
  geom_point(alpha=0.5)+
    #Add nice commas x axis
  scale_x_continuous(labels=dollar_format())+
  #Add nice commas y axis
scale_y_continuous(labels=comma_format())+
  #Add a linear regression line for all regions
  geom_smooth(aes(group=1),method="lm")+
  #Remove legend
  theme(legend.position = "none")+
   #Add title
  labs(title="Gross National Income and R downloads.",
       y="Downloads per country",
       x="Gross National Income")

```

```{r plot-three-HDI}
plot_grid(life_exp_plot, educ_plot, GNI_plot, rows = 3)
```


```{r four-var-fit}
four_fit <- lm(n ~ users + GNI + educ + life_exp, data=full_hdi)
summary(four_fit)$coef
```

Seems like separating the four doesn't lead to any significant coefficients beyond internet users. Interesting. I leave `HDI` in and it does, barely. I wonder what's behind the number of internet users... 

Which are the top 10 countreis 

```{r population-fit}
pop_fit <- lm(n ~ pop_2018 + HDI, data=full_hdi, na.action = na.omit)
summary(pop_fit)$coef
```

I added population as a variable, now since multicollinearity is a threat, that is adding variables that are highly correlated amongst themselves is as issue, I tested the correlation between internet users and population and it was `r cor(full_hdi$users,full_hdi$pop_2018, use = "pairwise.complete.obs")*100`. It's high. Because of this, I have to choose using either population or internet users. I fitted the model again with `HDI` and `pop_2018` and the adjusted R Squared is `r summary(pop_fit)$adj.r.squared*100`. With population instead of internet users, the model explains `r (summary(hdi_fit)$adj.r.squared-summary(pop_fit)$adj.r.squared)*100` less of the variation in the data.

So far, with all the variables I've tried so far, inclusion of the Human Development Index and internet users per country most explain variation in the data.

I wonder if I nested models by region... I wonder if that would change something. Interesting to try. I'll do this and then write it up.

```{r nesting-models}


nested_hdi <- full_hdi%>%
  group_by(region)%>%
  nest()
```


```{r plot-median-downloads-per-region, fig.cap="The median downloads per region was calculated using country-level data. The region with most median downloads is Europe, followed by Asia. The Americas, despite having the US as the country with most R Downloads in the world, is not the first since countries across the Americas do not have as many downloads."}
downloads_nested <- nested_hdi %>%
  mutate(median_n = map_dbl(data, ~median(.x$n)))

# Extract the median_n value by using unnest
downloads_nested %>% 
  unnest(median_n)%>%
  arrange(desc(median_n))%>%
  mutate(region=fct_reorder(region,median_n))%>%
  ggplot(aes(x=region, y=median_n, fill=region))+
  geom_col()+
  expand_limits(y=0)+
  coord_flip()+
  scale_y_continuous(labels=comma_format())+
  labs(title="median R Downloads by country per region",
       x="",
       y="R Downloads")+
  theme(legend.position = "none")
```


```{r linear-model-for-each-region}
regression_each_region <- nested_hdi %>%
  mutate(model=map(data, ~lm(formula=n~users, data=.x)))
```

```{r extracting-coefficients}
# Extract the coefficient statistics of each model into nested dataframes
model_coef_nested <- regression_each_region %>% 
    mutate(coef = map(model, ~broom::tidy(.x)))
    
# Simplify the coef dataframes for each model    
model_coef <- model_coef_nested %>%
    unnest(coef)
```

```{r table-significant-coefficient}
model_coef%>% 
  #Filter significant coefficient
  filter(p.value<0.05)%>%       
  arrange(p.value)%>%
  kable(caption="The model describing R downloads as a function of internet users per country and the HDI has significant coefficients for Asia and Europe.")
```

```{r assessing-model-fit}
# Extract the fit statistics of each model into dataframes
model_performance <- regression_each_region%>% 
    mutate(fit = map(model, ~glance(.x)))%>%
  unnest(fit)

```

```{r table-model-fit}
#Make a table with model fit in descending order #Filter significant coefficient
model_performance%>%
  arrange(desc(adj.r.squared))%>%
  select(region, adj.r.squared)%>%
  kable(caption="The model describing R downloads as a function of internet users per country and the HDI has significant coefficients for Asia and Europe.")
```


```{r augment-models}
models_augmented <- regression_each_region%>%
  mutate(augmented=map(model,~augment(.x)))%>%
  unnest(augmented)
```

```{r plot-model-fit}
models_augmented%>%
  ggplot(aes(x=users/1000, y=n))+
  geom_point(alpha=0.5)+
  geom_line(aes(y=.fitted), color="red", alpha=0.5)+
  facet_wrap(~region, scales = "free")+ 
    #Add nice commas x axis
scale_x_continuous(labels=comma_format())+
  #Add nice commas y axis
scale_y_continuous(labels=comma_format())+
labs(title="Top 10 countries by R downloads per 1,000 capita",
     subtitle = "Red dots represents fitted value, black dots are real values.",
       x="Divided by 1,000 users",
       y="Downloads")
```

## Cross validating the models with the MAP function Europe

```{r creating-train-test-data-europe}
set.seed(123)

full_europe <- full_hdi%>%
  filter(region=="Europe")
#Setting seed
downloads_split_europe <- initial_split(full_europe,prop=0.75)
#Store training_data
training_data_europe <-training(downloads_split_europe)
#Store testing_data
testing_data_europe <- testing(downloads_split_europe)
#Split the data into 3 folds
cv_split_europe <- vfold_cv(training_data_europe, v=3)
```

```{r creating-train-and-validate-data-from-each-fold-europe}
cv_data_europe <- cv_split_europe%>%
  #Create cross-validated data
  mutate(train=map(splits, ~training(.x)),
         validate=map(splits, ~testing(.x)))
```

```{r creating-models-with-cross-validated-data-europe}
cv_models_lm_europe <- cv_data_europe %>%
  #Create models
  mutate(model=map(train,~lm(formula=n ~ users, data = .x)))
```

## Measuring cross-validation performance

```{r extract-actual-and-predicted-values-europe}
cv_prep_lm_europe <- cv_models_lm_europe %>%
  #Create actual values
  mutate(validate_actual=map(validate, ~ .x$n),
  #Store predicted values
         validate_predicted=map2(model, validate,~predict(.x,.y)))
```

## Calculate performance

```{r calculate-MAE-europe}

cv_eval_lm_europe <- cv_prep_lm_europe%>%
  #Compute MAE
  mutate(validate_mae=map2_dbl(validate_actual,validate_predicted, ~mae(actual=.x,predicted = .y)))
```

## Cross validating the models with the MAP function Americas

```{r creating-train-test-data-americas}

full_americas <- full_hdi%>%
  filter(region=="Americas")
#Setting seed
downloads_split_americas <- initial_split(full_americas,prop=0.75)
#Store training_data
training_data_americas <-training(downloads_split_americas)
#Store testing_data
testing_data_americas <- testing(downloads_split_americas)
#Split the data into 3 folds
cv_split_americas <- vfold_cv(training_data_americas, v=3)
```

```{r creating-train-and-validate-data-from-each-fold-americas}
cv_data_americas <- cv_split_americas%>%
  #Create cross-validated data
  mutate(train=map(splits, ~training(.x)),
         validate=map(splits, ~testing(.x)))
```

```{r creating-models-with-cross-validated-data-americas}
cv_models_lm_americas <- cv_data_americas %>%
  #Create models
  mutate(model=map(train,~lm(formula=n ~ users, data = .x)))
```

## Measuring cross-validation performance

```{r extract-actual-and-predicted-values-americas}
cv_prep_lm_americas <- cv_models_lm_americas %>%
  #Create actual values
  mutate(validate_actual=map(validate, ~ .x$n),
         #Store predicted values
         validate_predicted=map2(model, validate,~predict(.x,.y)))
```

## Calculate performance

```{r calculate-MAE-americas}

cv_eval_lm_americas <- cv_prep_lm_americas%>%
  #Compute MAE
  mutate(validate_mae=map2_dbl(validate_actual,validate_predicted, ~mae(actual=.x,predicted = .y)))
```


## Cross validating the models with the MAP function Asia

```{r creating-train-test-data-asia}

full_asia <- full_hdi%>%
  filter(region=="Asia")
#Setting seed
downloads_split_asia <- initial_split(full_asia,prop=0.75)
#Store training_data
training_data_asia <-training(downloads_split_asia)
#Store testing_data
testing_data_asia <- testing(downloads_split_asia)
#Split the data into 3 folds
cv_split_asia <- vfold_cv(training_data_asia, v=3)
```

```{r creating-train-and-validate-data-from-each-fold-asia}
cv_data_asia <- cv_split_asia%>%
  #Create cross-validated data
  mutate(train=map(splits, ~training(.x)),
         validate=map(splits, ~testing(.x)))
```

```{r creating-models-with-cross-validated-data-asia}
cv_models_lm_asia <- cv_data_asia %>%
  #Create models
  mutate(model=map(train,~lm(formula=n ~ users, data = .x)))
```

## Measuring cross-validation performance

```{r extract-actual-and-predicted-values-asia}
cv_prep_lm_asia <- cv_models_lm_asia %>%
  #Create actual values
  mutate(validate_actual=map(validate, ~ .x$n),
         #Store predicted values
         validate_predicted=map2(model, validate,~predict(.x,.y)))
```

## Calculate performance

```{r calculate-MAE-asia}
cv_eval_lm_asia <- cv_prep_lm_asia%>%
  #Compute MAE
  mutate(validate_mae=map2_dbl(validate_actual,validate_predicted, ~mae(actual=.x,predicted = .y)))
```



## Testing the final model

```{r testing-the-final-model-europe}
#Store the best model with traning data
best_model_europe <- lm(n~users, data=training_data_europe)
#Store test data
test_actual_europe <- testing_data_europe$n
#Use the best model on the the test data
test_predict_europe <-predict(best_model_europe,testing_data_europe)
#Compute mae
mae(test_actual_europe,test_predict_europe)
```

```{r testing-the-final-model-americas}
#Store the best model with traning data
best_model_americas <- lm(n~users, data=training_data_americas)
#Store test data
test_actual_americas <- testing_data_americas$n
#Use the best model on the the test data
test_predict_americas <-predict(best_model_americas,testing_data_americas)
#Compute mae
mae(test_actual_americas,test_predict_americas)
```


```{r testing-the-final-model-asia}
#Store the best model with traning data
best_model_asia <- lm(n~users, data=training_data_asia)
#Store test data
test_actual_asia <- testing_data_asia$n
#Use the best model on the the test data
test_predict_asia <-predict(best_model_asia,testing_data_asia)
#Compute mae
mae(test_actual_asia,test_predict_asia)
```

# Reggression conclusions 

After trying different models, an aggregated model for the entire world, and the same model for each region, the variable most associated with R-Downloads was number of internet users. In Europe, Asia, and the Americas, users explains at least 75% of the variability in the data. For Africa, internet users only explained 26.8% of variation in the data. I'm excluding Oceania since it only had 8 observations. All the other regions had at least 30.

Despite explaining a lot of the variability in the model, after using cross-validation to test the final model, on average, our models deviate by `r mae(test_actual_europe,test_predict_europe)` for Europe, `r mae(test_actual_americas,test_predict_americas)` for the Americas, and `r mae(test_actual_asia,test_predict_asia)` for Asia.

# Onto maps

So now that we've seen how good (or bad our models can get), let's put that aside and see how we can plot downloads around the world and downloads per capita. I had taken a course and knew how to make chloropleth maps, a while ago. I have to freshen up and that's the aim of this session. 

```{r create-tibble-with-pop-and-users-only}
full_maps <- downloads_per_country%>%
  left_join(pop, by=c("name"="name"))%>%
  select(country, n, name, iso_3, region, sub_region, pop_2018)%>%
  mutate(n_capita=n/pop_2018)%>%
  filter(!is.na(n_capita),
         pop_2018!=0)
```

Excellent! Now that we have a tibble with not all countries, but at least 193 regions, we are ready to draw some maps I think!

```{r see-summary}
#view(dfSummary(world_full))
```

```{r read-map-as-sf-object}

map_path <- here::here("static","data","world_shapefile","world_shapefile.shp")
world <- sf::st_read(map_path)

```

```{r merge-full-maps-and-map-path}
world_full <- world%>%
  left_join(full_maps, by=c("ISO3"="iso_3"))
```

```{r plot-n-downloads-per-capita}
plot_downloads_per_capita <- tm_shape(world_full) + 
    tm_polygons(col="n_capita",n=4, style="jenks") + 
tm_layout("R Downloads per Capita in 2017/18", title.size=.8, frame = FALSE)+
  tm_scale_bar(breaks = c(0, 100, 200), size = 1) 

tmap_leaflet(plot_downloads_per_capita)
```


```{r plot-n-downloads}
plot_total_downloads <- tm_shape(world_full)+
    tm_polygons(col="n",n=4, style="jenks") + 
tm_layout("Total R Downloads in 2017/18", title.size=.8, frame = FALSE)+
  tm_scale_bar(breaks = c(0, 100, 200), size = 1) 

tmap_leaflet(plot_total_downloads)
```

