---
title: Rossman stores sales
author: Ángela D. Castillo-Gill
date: '2018-09-24'
slug: rossman-stores-sales
categories:
  - R
  - EDA
tags: 
  - Kaggle
  - Rossman
output:
  blogdown::html_page:
    toc: true
    number_sections: true
    toc_depth: 2
  fig_caption: true
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
x <-
  c("tidyverse",
    "knitr",
    "formatR",
    "stringr",
    "quantmode",
    "skimr",
    "lubridate",
    "tidyr",
    "formattable",
    "grid",
    "gridExtra",
    "kableExtra",
    "corrplot",
    "rms")

lapply(x, require, character.only = TRUE)


opts_chunk$set(echo = FALSE,
               warning = FALSE,
               error = FALSE,
               message = FALSE,
               collapse= TRUE,
               comment = NA,
               tidy = TRUE)
```

# Summary



# Purpose of this post


# The data

The data was obtained from the [Kaggle platform](https://www.kaggle.com/c/rossmann-store-sales/kernels?sortBy=hotness&group=everyone&pageSize=20&language=R&outputType=Data&competitionId=4594&kernelType=all).

```{r results='hide', message=FALSE}
path <- "/Volumes/TOSHIBAEXT/RStudio/blog/content/post/rossman_sales/"
#Read `trained_filled_gap.csv`
rossman <- read_csv(paste0(path,"train.csv"))
#Read `store.csv`
store <- read_csv(paste0(path,"store.csv"))
```

```{r}
#Making big dataset
full_store <- left_join(rossman,store,by="Store")
```


```{r}
#See dataset structure
str(full_store, give.attr = FALSE)
```

```{r}
#Removing NA's
full_store <- na.omit(full_store)
```

```{r}
# Visualization of correlations
full_store %>% select_if(is.numeric) %>%
  dplyr::select(-Store)%>%
  cor()%>%
  corrplot()
```

```{r}
#Making model
simple_model <- lm(Sales  ~ . - Store - StateHoliday - Promo2 - Date, 
                       data = full_store)
summary(simple_model)
```

```{r}
#Checking for multicollinearity
vif(simple_model)
```


```{r}
# Frequent days of week
ggplot(rossman) +
    geom_boxplot(aes(x = as.character(DayOfWeek), y = Sales))
```

```{r}
# Frequent days of week
ggplot(rossman) +
    geom_boxplot(aes(x = as.character(Promo), y = Sales))
```

```{r}
# Plot a histogram
ggplot(rossman) +
  geom_histogram(aes(x = Customers,
                     fill = factor(Open))) +
  facet_grid( ~ Open) + # Separate plots for Open = 1 vs. 0
  theme(legend.position = "none") # Don't show legend
```

```{r}
# Plot a histogram
ggplot(rossman) +
  geom_histogram(aes(x = Customers,
                     fill = factor(Promo))) +
  facet_grid( ~ Promo) + # Separate plots for Promo = 1 vs. 0
  theme(legend.position = "none") # Don't show legend
```

```{r}
# Plot a histogram
ggplot(rossman) +
  geom_histogram(aes(x = Customers,
                     fill = factor(StateHoliday))) +
  facet_grid( ~ StateHoliday) + # Separate plots for StateHoliday = 1 vs. 0
  theme(legend.position = "none") # Don't show legend
```

```{r}
# Plot a histogram
ggplot(rossman) +
  geom_histogram(aes(x = Customers,
                     fill = factor(SchoolHoliday))) +
  facet_grid( ~ SchoolHoliday) + # Separate plots for SchoolHoliday = 1 vs. 0
  theme(legend.position = "none") # Don't show legend
```


The codebook was included in the Kaggle description: 

- `customerID`: "Customer ID"
- `gender`: "The amount disbursed by the field agent to the borrower(USD)"
- `SeniorCitizen`: "More granular category"
- `sector`: "High level category"
- `use`: "Exact usage of loan amount"
- `country_code`: "ISO country code of country in which loan was disbursed"
- `country`: "Full country name of country in which loan was disbursed"
- `region`: "Full region name within the country"
- `currency`: "The currency in which the loan was disbursed"
- `partner_id`: "ID of partner organization"



# Data cleaning 
```{r}
#Separate Date into Year Month Day

rossman_date <- cbind(rossman, year=year(rossman$Date),
                      month=month(rossman$Date),
                      day=day(rossman$Date))

```

From  at the structure of our data, we can soon see there are a few bits that don't make sense and need to be fixed. 

**a. Missing values**




# Exploratory data analysis



# Techniques used



# Questions from this analysis

```{r}
#Calculating annual sales and customers per per store
annual_sales <-  aggregate (Sales ~ year + Store,
                            data = rossman_date,
                            sum)%>%
  spread(key=year,
         value = Sales)

#Changing names to R friendly names

colnames(annual_sales) <- c("Store",
                            "sales_2013",
                            "sales_2014",
                            "sales_2015")
```

```{r}
#Joining store characteristics with sales data
full_store <- left_join(annual_sales,store,by="Store")
```


```{r}
#Checking linear relationship
ggplot(annual_sales, aes(x=sales_2013,
                         y=sales_2014))+geom_point()

```

```{r}
#Predicting sales one year based on the previous years
salesSimpleModel <- lm(sales_2014 ~ sales_2013, 
                       data = annual_sales)
```

```{r}
#Looking at the distribution of each variable
ggplot(full_store,aes(x=CompetitionDistance,y=sales_2015))+
  geom_point()
```

```{r}

```


```{r}

```

