---
title: Exploring Kiva loans
author: Ángela D. Castillo-Gill
date: '2018-07-07'
slug: exploring-kiva-loans
categories:
  - R
  - EDA
tags: 
  - Kaggle
  - Kiva
summary: What do Kiva loans around the world look like?
output:
  blogdown::html_page:
    toc: true
    number_sections: true
    toc_depth: 2
  fig_caption: true
---


```{r setup, include=FALSE}
x <-
  c("tidyverse",
    "knitr",
    "formatR",
    "stringr",
    "quantmode",
    "skimr",
    "lubridate",
    "tidyr",
    "formattable",
    "grid",
    "gridExtra",
    "kableExtra")

lapply(x, require, character.only = TRUE)


opts_chunk$set(echo = FALSE,
               warning = FALSE,
               error = FALSE,
               message = FALSE,
               collapse= TRUE,
               comment = NA,
               tidy = TRUE)

```

# Summary


# Purpose of this post

Muhammad Yunus and the Grameen Bank won the [Nobel Peace Prize](https://www.nobelpeaceprize.org/Prize-winners/Winners/2006) in 2006 for "their efforts through microcredit to create economic and social development from below."

Back in 1976, Yunus, a professor at the University of Chittagong (Bangladesh) at the time, [noticed that small amounts of money could make a substantial difference to people living in poverty](https://www.bbc.co.uk/news/world-south-asia-11901625). He started to loan money to people that didn't meet the requirements listed by the mainstream banking system. Not only did these type of loans turn out to be effective to ["emerge" from poverty](https://www.nobelpeaceprize.org/Prize-winners/Prizewinner-documentation/Muhammad-Yunus-Grameen-Bank), but default rates were lower than those of commercial banks, [reported at 2%](https://www.gdrc.org/icm/grameen-article4.html). Eventually, in October 1983, Muhammad Yunus founded [Grameen Bank](https://grameenfoundation.org/about/history), considered to be the first microfinance institution. 

Founded in 2005, [Kiva](www.kiva.org), has the same mission as Grameen Bank except that anyone can become a Kiva banker. This online platform enables microcredit lending to help low-income entrepreneurs around the world with a couple of cliks. Pretty neat, huh? In this post, I unpack a large dataset published by [Kiva on the Kaggle platform](https://www.kaggle.com/kiva/data-science-for-good-kiva-crowdfunding) and explore these cool microloans.

# The data

The dataset was published during the first months of 2018. 

# What I did

If you would like to have a look at all the code used in this post, feel free to visit my [kernel on kaggle in R Markdown format](https://www.kaggle.com/adcastillogill/exploring-kiva-loans). 

The first step was to have a look at the dataset. This is pretty exciting. The zip file (232.7 MB) from Kaggle contained four files: `kiva_loans.csv`, `kiva_mpi_region_locations.csv`,`loan_theme_ids.csv`, and `loan_themes_by_region.csv`. 

```{r results='hide', message=FALSE}
path <- "/Volumes/TOSHIBAEXT/RStudio/blog/content/post/kiva_data/"
#Read `kiva_loans.csv`
kiva_loans <- read_csv(paste0(path,"kiva_loans.csv"))

```

Let's see what's inside. 

```{r}
#Get a glimpse of kiva_loans
glimpse(kiva_loans)
```

Wow! 671,205 observations and 20 variables.

Let's have a closer look at the data and see what each variable looks like.

- `funded_amount`: "The amount disbursed by Kiva to the field agent(USD)"
- `loan_amount`: "The amount disbursed by the field agent to the borrower(USD)"

Here, I am interested in the latter of those two variables as my response variable. 

- `activity`: "More granular category"
- `sector`: "High level category"
- `use`: "Exact usage of loan amount"

From these, I am going to pick `sector` because at first, I want to see if I can treat it like a factor in our regressions. The other ones are perhaps to specific to create levels.

- `country_code`: "ISO country code of country in which loan was disbursed"
- `country`: "Full country name of country in which loan was disbursed"
- `region`: "Full region name within the country"

I'm going to stick with `country_code` to make a factor with each country as a level.

- `currency`: "The currency in which the loan was disbursed"

Super important! In order to compare all loans, we need to translate them into the same currency.

- `partner_id`: "ID of partner organization"

This variable has a lot of missing values and the Kiva explanation on Kaggle doesn't go much further. For now, I will exclude `partner_id`. 

- `posted_time`: "The time at which the loan is posted on Kiva by the field agent"
- `disbursed_time`: "The time at which the loan is disbursed by the field agent to the borrower"
- `funded_time`: "The time at which the loan posted to Kiva gets funded by lenders completely"
- `term_in_months`: "The duration for which the loan was disbursed in months"
- `lender_count`: "The total number of lenders that contributed to this loan"
- `borrower_genders`: "Comma separated M,F letters, where each instance represents a single male/female in the group"
- `repayment_interval`: Not specified so we'll assume that it means the standard definition of when the loan is repaid back to the lender.

I'm including all in this last batch, they provide specific information for each loan which may prove useful later. Now let's create a new dataset with the selected variables with a few changes. 

# Data cleaning 
```{r}

#Select variables 
kiva_loans %>%
  dplyr::select (id,
          loan_amount,
          sector,
          country_code,
          currency,
          posted_time,
          disbursed_time,
          term_in_months,
          lender_count,
          borrower_genders,
          repayment_interval,
          funded_time) -> kiva_loans

#Convert country_code, sector, currency, borrower_genders, and repayment_interval to factors.
categorical <- c("sector",
                 "country_code",
                 "currency",
                 "borrower_genders",
                 "repayment_interval")

kiva_loans[categorical] <- lapply(kiva_loans[categorical],
                                  factor)
```

From  at the structure of our data, we can soon see there are a few bits that don't make sense and need to be fixed. 


**a. Missing values**

Missing values, aka `NA's`, we could simply remove them since we have plenty of observations. But what if by removing them we introduce some bias. For example, `borrower_genders` has plenty of missing values, but what if this could say something about certain types of loans. For now we will leave them in.

**b.`borrower_genders`**

From the variable descriptions, we expected `borrower_genders`to have only two levels, `male`or `female`. Here we see there are many more levels, 11,298 to be precise. We know something is wrong here. So we'll fix that first. 

```{r }

#Turn kiva_loans back into an atomic vector of characters
kiva_loans$borrower_genders <- as.character(kiva_loans$borrower_genders)

#Split each row into multiple strings
genders <- str_split(kiva_loans$borrower_genders, ", ")

#Assign 1 for female and 0 for male
females_count <- lapply(genders, str_count, pattern="female")

#Count how many females are in each list element
females_sum <- lapply(females_count, sum)

#Create a function that takes the two lists just made
compare_genders <- function(females_count, females_sum){
  
  #Set "single_female" when female sum and length is 1
  ifelse (females_sum==1&&length(females_count)==1, "single_female",
          
  #Set "mult_females" when female sum and length females_count is the same and different to zero
  ifelse(females_sum==length(females_count)&&females_sum!=0,"mult_females",
         
  #Set "single_male" #Set "mult_males" when female sum is zero and the count of individuals is 1
  ifelse(females_sum==0&&length(females_count)==1,"single_male",
         
  #Set "mult_males" when female sum is zero and the count of individuals is greater than 1
  ifelse(females_sum==0&&length(females_count)>1,"mult_males",
         
  #Set "mixed_genders" when the sum of females is different to the count of individual       
  ifelse(females_sum!=length(females_count),"mixed_genders","strange")))))
  
}
#Apply function compare_genders
kiva_loans$borrower_genders <- mapply(compare_genders,females_count,females_sum)
```

Let's see if all is well...
```{r}

#Make kiva_loans$borrower_genders a factor again
kiva_loans$borrower_genders <- as.factor(kiva_loans$borrower_genders)

#Check levels
levels(kiva_loans$borrower_genders)
```
Nothing `strange`! Five levels as we wanted - fabulous! Onwards. 

**c.`loan_amounts`**

Now, since we're trying to predict `loan_amounts`, it's better if they're all in the same currency. Let's translate them to USD.

```{r eval=FALSE}

for (curr in levels(kiva_loans$currency))
  
  #Check that all currencies do not return any errors when requesting a currency conversion
  tryCatch(quantmod::getQuote(paste0(curr, "=X")), error = print)
```

It seems like two currencies are returning an error. Let's filter them out. 

```{r}

#Turning kiva_loans back into a character vector
kiva_loans$currency <- as.character(kiva_loans$currency)

kiva_loans%>%
  #Filter out "SSP"
  dplyr::filter(currency!="SSP")%>% 
  
  #Filter out "ZWD"
  filter(currency!="ZWD")->kiva_loans 

#Turn kiva_loans$currency back into a factor
kiva_loans$currency <- as.factor(kiva_loans$currency) 

#Store all levels as vector `from`
from <- as.vector(levels(kiva_loans$currency)) 

#Create a vector `to` with requested currency "USD"
to <- rep(length(levels(kiva_loans$currency)),x="USD") 

#Create a dataframe `currencies` with conversion rates `Last`
currencies <- quantmod::getQuote(paste0(from, to, "=X"))
  

```

```{r}
#Set row names as columns in the currencies dataframe

currencies <- rownames_to_column(currencies)

#Split currencies$rownames by extracting the first 3 characters

currencies$currency <-  substring(currencies$rowname, 1,3)

#currencies <- readRDS(file="/Users/AngelaDelgado/Desktop/currencies.rds")

kiva_loans%>%
  #Join currencies to kiva_loans by="currency"
  left_join(currencies, by="currency") -> kiva_loans

kiva_loans%>%
  #Convert the loan from its original currency to USD
  mutate(loan_usd=loan_amount*Last)-> kiva_loans
```


**d.`country_codes`**

Finally, with 86 countries, we have 86 levels. Perhaps it would be interesting to create another category called continent to produce less levels and have a better understanding of the overall function of regional distributions.

Our country codes are in the ISO3166 format, so we will use the associated continent code found [here](https://dev.maxmind.com/geoip/legacy/codes/country_continent/).

```{r}
#Read `country_continent.csv`
continents <- read_csv(paste(path,"country_continent.csv",sep = ""))
```

We will do what we previously did for currencies, which is to join both tables. 

```{r}
kiva_loans%>%
  #Join currencies to kiva_loans by="country_code"
  left_join(continents, by="country_code") -> kiva_loans

```

Let's see what happened...
```{r}
kiva_loans$continent <- as.factor(kiva_loans$continent)
levels(kiva_loans$continent)
```
5 continents! Africa, Asia, Europe, Oceania, and South America. Granted, we've lost spacial resolution. However, for our initial purpose, this is exactly what we wanted. 

**e. Dates**

Dates aren't numeric, but we can make them. Let's calculate two lengths of time that I think are interesting. 
First, how much time passes from the moment the loan is posted to the moment it's disbursed (`total_time`). Second, how long does a loan take to get funded (`giving_time`)? 

```{r total_time_creation}
kiva_loans%>%
  mutate(total_time_num=as.numeric(difftime(posted_time,
                                            disbursed_time,
                                            units = "days")),
         giving_time_num=as.numeric(difftime(funded_time,
                                             posted_time,
                                             units = "days"))) ->kiva_loans

```


```{r, results='hide'}
#Select variables 
kiva_loans %>%
  dplyr::select (id,
          loan_usd,
          sector, 
          continent,
          term_in_months,
          lender_count,
          borrower_genders,
          repayment_interval,
          total_time_num,
          giving_time_num,
          country_code,
          posted_time,
          funded_time,
          disbursed_time)%>%
  arrange(posted_time)-> kiva_loans
kiva_loans$country_code <- as.factor(kiva_loans$country_code)
kiva_loans$id <- as.character(kiva_loans$id)
#Skim the dataset
skim(kiva_loans)
```


# Exploratory data analysis

Now let's describe the data. We could jump into the regression trees but exploring the data will help us interpret our predictions better. Do some extra work now and enjoy the benefits later. 


Let's make some tables to understand our factors.

```{r}
prop.table(table(kiva_loans$borrower_genders))%>%
  as.tibble()%>%
  mutate(n=n*100)%>%
  arrange(desc(n))%>%
  kable(caption="A single female is the most common type of borrower gender with over half of all Kiva loans requested.",
        col.names = c("Gender","Percentage")) %>%
 kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
               full_width = F)



prop.table(table(kiva_loans$continent))%>%
  as.tibble()%>%
  mutate(n=n*100)%>%
  arrange(desc(n))%>%
  kable(caption="Half of all Kiva loans are requested in Asia followed by Africa, South America, Oceania and finally the European Union.",
        col.names = c("Continents","Percentage")) %>%
 kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
               full_width = F)



prop.table(table(kiva_loans$repayment_interval))%>%
  as.tibble()%>%
  mutate(n=n*100)%>%
  arrange(desc(n))%>%
  kable(caption="The most popular type of repayment interval is monthly.",
        col.names = c("Repayment interval","Percentage")) %>%
 kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
               full_width = F)



prop.table(table(kiva_loans$sector))%>%
  as.tibble()%>%
  mutate(n=n*100)%>%
  arrange(desc(n))%>%
  kable(caption="Generally, the most frequent use of loans is agriculture, followed by food and retail.",
        col.names = c("Sector","Percentage")) %>%
 kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
               full_width = F)


```

Notice most loans are requested by single females, that the least amount of loans are given in the EU, that weekly repayment is an unpopular form of paying the loan back and that entertainment, wholesale, manufacturing, and construction amount less than 2.2% of sectors. Let's remove these categories that are not representative of the data. 

```{r}
kiva_loans <-  kiva_loans%>%
  filter(continent!="EU",
         repayment_interval!="weekly",
         sector!="Construction"&
          sector!="Entertainment"&
           sector!= "Manufacturing"&
            sector!= "Wholesale")%>%
  droplevels()
```

Ideal. Now let's look at our factors in an easier way to parse. Let's compare genders by continents to see if we can spot anything worth noting.

```{r continent-and-borrower-genders-plots, fig.cap="Four continents in descending order from the continent that requests the most loans (Asia). Notice Europe and North America do not appear. `mult_females` or `mult_males` means more than one female or male.",fig.width=15}
AF_borrower_genders <- na.omit(kiva_loans) %>%
  filter(continent == "AF") %>%
  dplyr::count(borrower_genders) %>%
  mutate(borrower_genders = reorder(borrower_genders, n)) %>%
  ggplot(aes(
    x = borrower_genders,
    y = n,
    fill = borrower_genders,
    color = borrower_genders
  )) +
  geom_col(show.legend = FALSE, alpha = 0.7) +
  theme_minimal() +
  labs(title = "Africa",
       x = "",
       y = "Number of loans") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.ticks.x = element_blank()) +
  #guides(fill = guide_legend(title = "Borrower genders")) +
  scale_fill_viridis_d() +
  scale_color_viridis_d() +
  coord_flip(ylim = c(0,250000))


OC_borrower_genders <- na.omit(kiva_loans) %>%
  filter(continent == "OC") %>%
  dplyr::count(borrower_genders) %>%
  mutate(borrower_genders = reorder(borrower_genders, n)) %>%
  ggplot(aes(
    x = borrower_genders,
    y = n,
    fill = borrower_genders,
    color = borrower_genders
  )) +
  geom_col(show.legend = FALSE, alpha = 0.7) +
  theme_minimal() +
  labs(title = "Oceania",
       x = "",
       y = "Number of loans") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.ticks.x = element_blank()) +
  #guides(fill = guide_legend(title = "Borrower genders")) +
  scale_fill_viridis_d() +
  scale_color_viridis_d() +
  coord_flip(ylim = c(0,250000))


AS_borrower_genders <- na.omit(kiva_loans) %>%
  filter(continent == "AS") %>%
  dplyr::count(borrower_genders) %>%
  mutate(borrower_genders = reorder(borrower_genders, n)) %>%
  ggplot(aes(
    x = borrower_genders,
    y = n,
    fill = borrower_genders,
    color = borrower_genders
  )) +
  geom_col(show.legend = FALSE, alpha = 0.7) +
  theme_minimal() +
  labs(title = "Asia",
       x = "",
       y = "Number of loans") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.ticks.x = element_blank()) +
  #guides(fill = guide_legend(title = "Borrower genders")) +
  scale_fill_viridis_d() +
  scale_color_viridis_d() +
  coord_flip(ylim = c(0,250000))


SA_borrower_genders <- na.omit(kiva_loans) %>%
  filter(continent == "SA") %>%
  dplyr::count(borrower_genders) %>%
  mutate(borrower_genders = reorder(borrower_genders, n)) %>%
  ggplot(aes(
    x = borrower_genders,
    y = n,
    fill = borrower_genders,
    color = borrower_genders
  )) +
  geom_col(show.legend = FALSE, alpha = 0.7) +
  theme_minimal() +
  labs(title = "South America",
       x = "",
       y = "Number of loans") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.ticks.x = element_blank()) +
  #guides(fill = guide_legend(title = "Borrower genders")) +
  scale_fill_viridis_d() +
  scale_color_viridis_d() +
  coord_flip(ylim = c(0,250000))

                 

grid.arrange(
  AS_borrower_genders,
  AF_borrower_genders,
  SA_borrower_genders,
  OC_borrower_genders,
  nrow = 1,
  top = grid::textGrob("Borrower genders by continent",
                       x=0,
                       hjust=0,
                       gp=gpar(fontsize=30,
                         font=8))
)
```

In all continents, single females request the most loans followed by single males. In Asia and South America, the third category is multiple females while in Africa it's mixed genders. 

```{r repayment-interval-and-continent, fig.cap="Most popular repayment styles by continent",fig.width=15}
AF_repayment_interval <- kiva_loans %>%
  filter(continent=="AF")%>%
  dplyr::count(repayment_interval)%>%
  mutate(repayment_interval=reorder(repayment_interval, n))%>%
  ggplot(aes(
    x = repayment_interval,
    y = n,
    fill = repayment_interval,
    color = repayment_interval)) +
  geom_col(show.legend = FALSE, alpha=0.7) +
  theme_minimal() +
  labs(title = "Africa",
       x = "",
       y = "Number of loans") +
  theme(axis.text.x = element_text(angle = 45, hjust=1),
        axis.ticks.x = element_blank()) +
  #guides(fill = guide_legend(title = "Repayment interval")) +
  scale_fill_viridis_d()+
  scale_color_viridis_d()+
  coord_flip(ylim = c(0,170000))


OC_repayment_interval <- kiva_loans %>%
  filter(continent=="OC")%>%
  dplyr::count(repayment_interval)%>%
  mutate(repayment_interval=reorder(repayment_interval, n))%>%
  ggplot(aes(
    x = repayment_interval,
    y = n,
    fill = repayment_interval,
    color = repayment_interval)) +
  geom_col(show.legend = FALSE, alpha=0.7) +
  theme_minimal() +
  labs(title = "Oceania",
       x = "",
       y = "Number of loans") +
  theme(axis.text.x = element_text(angle = 45, hjust=1),
        axis.ticks.x = element_blank()) +
  #guides(fill = guide_legend(title = "Repayment interval")) +
  scale_fill_viridis_d()+
  scale_color_viridis_d()+
  coord_flip(ylim = c(0,170000))


AS_repayment_interval <- kiva_loans %>%
  filter(continent=="AS")%>%
  dplyr::count(repayment_interval)%>%
  mutate(repayment_interval=reorder(repayment_interval, n))%>%
  ggplot(aes(
    x = repayment_interval,
    y = n,
    fill = repayment_interval,
    color = repayment_interval)) +
  geom_col(show.legend = FALSE, alpha=0.7) +
  theme_minimal() +
  labs(title = "Asia",
       x = "",
       y = "Number of loans") +
  theme(axis.text.x = element_text(angle = 45, hjust=1),
        axis.ticks.x = element_blank()) +
  #guides(fill = guide_legend(title = "Repayment interval")) +
  scale_fill_viridis_d()+
  scale_color_viridis_d()+
  coord_flip(ylim = c(0,170000))


SA_repayment_interval <- kiva_loans %>%
  filter(continent=="SA")%>%
  dplyr::count(repayment_interval)%>%
  mutate(repayment_interval=reorder(repayment_interval, n))%>%
  ggplot(aes(
    x = repayment_interval,
    y = n,
    fill = repayment_interval,
    color = repayment_interval)) +
  geom_col(show.legend = FALSE, alpha=0.7) +
  theme_minimal() +
  labs(title = "South America",
       x = "",
       y = "Number of loans") +
  theme(axis.text.x = element_text(angle = 45, hjust=1),
        axis.ticks.x = element_blank()) +
  #guides(fill = guide_legend(title = "Repayment interval")) +
  scale_fill_viridis_d()+
  scale_color_viridis_d()+
  coord_flip(ylim = c(0,170000))

grid.arrange(AS_repayment_interval,
             AF_repayment_interval,
             SA_repayment_interval, OC_repayment_interval,
             nrow=1,  
             top = grid::textGrob("Repayment interval by continent",
                       x=0,
                       hjust=0,
                       gp=gpar(fontsize=30,
                               font=8))
)

```

In Asia and Oceania, the most popular repayment interval is irregular while in Africa and South America it's monthly. 

```{r sectors-and-continents, fig.cap="Most popular uses of loans by continent",fig.width=12}
AF_sector <- kiva_loans %>%
filter(continent=="AF")%>%
dplyr::count(sector)%>%
mutate(sector=reorder(sector, n))%>%
ggplot(aes(
x = sector,
y = n,
fill = sector,
color = sector)) +
geom_col(show.legend = FALSE, alpha=0.7) +
theme_minimal() +
labs(title = "Africa",
x = "",
y = "Number of loans") +
theme(axis.text.x = element_text(angle = 45, hjust=1),
axis.ticks.x = element_blank()) +
#guides(fill = guide_legend(title = "Sector")) +
scale_fill_viridis_d()+
scale_color_viridis_d()+
coord_flip()


OC_sector <- kiva_loans %>%
filter(continent=="OC")%>%
dplyr::count(sector)%>%
mutate(sector=reorder(sector, n))%>%
ggplot(aes(
x = sector,
y = n,
fill = sector,
color = sector)) +
geom_col(show.legend = FALSE, alpha=0.7) +
theme_minimal() +
labs(title = "Oceania",
x = "",
y = "Number of loans") +
theme(axis.text.x = element_text(angle = 45, hjust=1),
axis.ticks.x = element_blank()) +
#guides(fill = guide_legend(title = "Sector")) +
scale_fill_viridis_d()+
scale_color_viridis_d()+
coord_flip()


AS_sector <- kiva_loans %>%
filter(continent=="AS")%>%
dplyr::count(sector)%>%
mutate(sector=reorder(sector, n))%>%
ggplot(aes(
x = sector,
y = n,
fill = sector,
color = sector)) +
geom_col(show.legend = FALSE, alpha=0.7) +
theme_minimal() +
labs(title = "Asia",
x = "",
y = "Number of loans") +
theme(axis.text.x = element_text(angle = 45, hjust=1),
axis.ticks.x = element_blank()) +
#guides(fill = guide_legend(title = "Sector")) +
scale_fill_viridis_d()+
scale_color_viridis_d()+
coord_flip()


SA_sector <- kiva_loans %>%
filter(continent=="SA")%>%
dplyr::count(sector)%>%
mutate(sector=reorder(sector, n))%>%
ggplot(aes(
x = sector,
y = n,
fill = sector,
color = sector)) +
geom_col(show.legend = FALSE, alpha=0.7) +
theme_minimal() +
labs(title = "South America",
x = "",
y = "Number of loans") +
theme(axis.text.x = element_text(angle = 45, hjust=1),
axis.ticks.x = element_blank()) +
#guides(fill = guide_legend(title = "Sector")) +
scale_fill_viridis_d()+
scale_color_viridis_d()+
coord_flip()

grid.arrange(AS_sector,
             AF_sector,
             SA_sector,
             OC_sector,
             nrow=1,
             top = grid::textGrob("Loan use by continent",
                       x=0,
                       hjust=0,
                       gp=gpar(fontsize=30,
                               font=8))
)
```

In every continent, the three most popular sectors are agriculture, retail, and food. 

Finally, we still would like to understand which countries are requesting the most loans. Let'see.

```{r countries-plot, fig.cap="Top 10 countries requesting Kiva loans"}
kiva_loans %>%
  group_by(country_code) %>%
  dplyr::count(country_code, sort = TRUE) %>%
  ungroup() %>%
  top_n(10) %>%
  ggplot(aes(reorder(x = country_code, n), y = n)) +
  geom_col(position = "dodge",
  color = "purple",
  fill = "purple",
  alpha=0.6) +
  coord_flip() +
  theme_minimal() +
  labs(title = "Top 10 countries requesting Kiva loans",
  x = "",
  y = "Number of loans")
```

We see that the Philippines is the country that requests the mosts loans with over double as the second country Kenya `KE`. Other top countries Cambodia `KH`, Pakistan `PK`, Peru `PE`, Colombia `CO`, Uganda `UG`, Tajikistan `TJ`, Ecuador `EC`, and Paraguay `PY`.


Now let's look at the distributions of our numeric variables.

```{r loan-usd-distribution, fig.cap="Distribution of Kiva loans in USD"}
ggplot(kiva_loans, aes(x = loan_usd)) +
  geom_density() +
  theme_classic() +
  scale_x_continuous(limits = c(0, 15)) +
  geom_vline(
  xintercept = median(kiva_loans$loan_usd),
  color = "red",
  linetype = "dashed") +
  annotate(
  "text",
  label = paste("Median =",
  round(median(kiva_loans$loan_usd), digits = 2), "USD"),
  x = 10,
  y = 0.2,
  color = "red"
  ) +
  labs(title = "Distribution of Kiva loans' amounts in USD",
  x = "US Dollars",
  y = "Density")
```

Figure \@ref(fig:loan-usd-distribution), shows that the distribution of Kiva loans in US dollars is right skewed and multimodal with its biggest peak around $1 and its second around $4. The median is about `r round(median(kiva_loans$loan_usd),digits=2)`. 

```{r term_in_months_distribution}
ggplot(kiva_loans,
       aes(x=term_in_months))+
  geom_density()+
  theme_classic()+
  scale_x_continuous(limits=c(0,50))+
  geom_vline(xintercept = median(kiva_loans$term_in_months),
             color="red",
             linetype="dashed")+
  annotate("text",
           label=paste("Median =",                          median(kiva_loans$term_in_months),"months"),
                x=18,
                y=0.2,
           color="red")+
    labs(title = "Distribution of Kiva loans' terms in months",
  x = "Months",
  y = "Density")
```
The loan term in months is a bimodal distribution with its first peak around 8 months and its second around 12, the median. 
```{r lender_count}
ggplot(kiva_loans,
       aes(x=lender_count))+
  geom_histogram()+
   theme_classic()+
  scale_x_continuous(limits=c(0,100))+
  geom_vline(xintercept = median(kiva_loans$lender_count),
             color="red",
             linetype="dashed")+
  annotate("text",
           label=paste("Median=",
                              median(kiva_loans$lender_count),"lenders"),
                x=21,
                y=150000,
           color="red")+
    labs(title = "Distribution of number of lenders per Kiva loan",
  x = "Number of lenders",
  y = "Frequency")
```
The lender count has a text book right-skewed distribution with a median of 12 lenders. 
```{r giving_time}
ggplot(kiva_loans,
       aes(x=giving_time_num))+
  geom_density()+
   theme_classic()+
  scale_x_continuous(limits=c(0,50))+
  geom_vline(xintercept = median(kiva_loans$giving_time_num,na.rm=TRUE),
             color="red",
             linetype="dashed")+
  annotate("text",
           label=paste("Median=",
                       round(median(kiva_loans$giving_time_num,na.rm = TRUE)),digits=2,"days"),
           x=median(kiva_loans$giving_time_num,na.rm=TRUE)+5,
           y=0.1,
           color="red")+
    labs(title = "Distribution of time to fully fund a Kiva loan",
  x = "Days",
  y = "Density")
```
The median time between posting the loan and the loan being fully funded is about 9.2 days. It is a bimodal distribution with its first peak around 6 days and its second peak around 30 days. 
```{r total_time}
ggplot(kiva_loans,
       aes(x=total_time_num))+
  geom_density()+
   theme_classic()+
  scale_x_continuous(limits=c(0,50))+
  geom_vline(xintercept = median(kiva_loans$total_time_num,
                                 na.rm=TRUE),
             color="red",
             linetype="dashed")+
  annotate("text",
           label=paste("Median=",
                       round(median(kiva_loans$total_time_num,
                                    na.rm = TRUE)),digits=2,"days"),
           x=15+5,
           y=0.1,
           color="red")+
    labs(title = "Distribution of time between posting and disbursing a Kiva loan",
  x = "Months",
  y = "Density")
```
The total time between posting the loan and it being disbursed has a funky looking multimodal distribution. It seems to be cut off at 30 days with peaks at 7, 14, 21 and 30 days. Median time is 17.2 days. 


# Skills practiced in this post 

- The `quantmode` package was new to me. I used this to convert all loans into a unique currency (US dollars) for comparison. There were two currencies that were unavailable and because I hadn't seen the warnings, the whole chunk kept throwing an error that was a bit frustrating. 

- Taking all the different levels that came in the `borrower_genders` variable and creating five neat categories to better understand who are the borrowers was good practice with lists and the `stringr` package. 


# What I'm curious about

- Why is `Retail` and not `Food` (as in other regions) the second most common use for loans in Asia. 

- Who are the givers? Where are they? Does proximity of the lender to the borrower have anything to do with funding times?

- Does the Kiva website have anything to do with funding times? For example, `giving_time`, the time between posting the loan and the loan being fully funded, has two peaks, at around one week and one month. Is this due to the platform and the promotion of loans that have been posted for a certain amount of time?
