---
title: Rossman stores sales
author: Ángela D. Castillo-Gill
date: '2018-09-24'
slug: rossman-stores-sales
categories:
  - R
  - EDA
tags: 
  - Kaggle
  - Rossman
summary: Which stores are the most profitable?
output:
  blogdown::html_page:
    toc: true
    number_sections: true
    toc_depth: 2
  fig_caption: true
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
x <-
  c("tidyverse",
    "knitr",
    "formatR",
    "stringr",
    "quantmode",
    "skimr",
    "lubridate",
    "tidyr",
    "formattable",
    "grid",
    "gridExtra",
    "kableExtra",
    "corrplot",
    "rms")

lapply(x, require, character.only = TRUE)


opts_chunk$set(echo = FALSE,
               warning = FALSE,
               error = FALSE,
               message = FALSE,
               collapse= TRUE,
               comment = NA,
               tidy = TRUE)
```

# Summary



# Purpose of this post


# The data



```{r results='hide', message=FALSE}
path <- "/Volumes/TOSHIBAEXT/RStudio/blog/content/post/rossman_sales/"
#Read `trained_filled_gap.csv`
rossman <- read_csv(paste0(path,"train.csv"))
#Read `store.csv`
store <- read_csv(paste0(path,"store.csv"))
```

```{r}
#Make big dataset and omit NA's
full_store <- na.omit(rossman)%>%
  left_join(store,by="Store")
```

The data was obtained from the [Kaggle platform](https://www.kaggle.com/c/rossmann-store-sales/kernels?sortBy=hotness&group=everyone&pageSize=20&language=R&outputType=Data&competitionId=4594&kernelType=all). After sales data with unique store data, the final data set is 986,159 observations with 18 variables. Each observation correponds to the turnover for a given day.

```{r}
#See dataset structure
str(full_store, give.attr = FALSE)
```



**Codebook**

The codebook was included in the Kaggle description but I'm including them here for clarity. 

- **Store**: A unique I for each store
- **DayOfWeek**: Day of the week
- **Date**: Date
- **Sales**: Turnover for any given day
- **Customers**: The number of customers on a given day
- **Open**: An indicator for whether the store was open (0 for closed, 1 for open)
- **Promo**: Indicates whether a store is running a promo on that day
- **StateHoliday**: Indicates if the store, on a day, was open or closed
- **SchoolHoliday**: Indicates if the store, on a day, was open or closed
- **StoreType**: Differentiates between four differente store models: a, b, c, and d.
- **Assortment**: Describes an assortment level: a = basic, b = extra, c = extended
- **CompetitionDistance**: Distance in meters to the nearest competitor store
- **CompetitionOpenSince [month/year]**: Gives the approximate year and month of the time the nearest competitor was opened
- **Promo**: Indicates whether a store is running a promo on that day
- **Promo2**: A continuing and consecutive promotion for some stores (0 not participating, 1 yes)
- **Promo2Since [year/week]** - Describes the year and calendar week when the store started participating in Promo2
- **PromoInterval**: Describes the consecutive intervals Promo2 started, naming the months the promotion is started anew. E.g. "Feb,May,Aug,Nov" means each round starts in February, May, August, November of any given year for that store

```{r declaring variables}

#Declare factors
cols <-c("DayOfWeek",
         "Open",
         "Promo",
         "StateHoliday",
         "SchoolHoliday",
         "StoreType",
         "Assortment",
         "Promo2")

#Make factors
full_store[cols] <- lapply(full_store[cols], factor)
```



```{r}
# # Visualization of correlations
# full_store %>% select_if(is.numeric) %>%
#   dplyr::select(-Store)%>%
#   cor()%>%
#   corrplot()
```



```{r}
#Checking for multicollinearity
#vif(simple_model)
```


```{r}
# Frequent days of week
# ggplot(rossman) +
#     geom_boxplot(aes(x = as.character(DayOfWeek), y = Sales))
```

```{r}
# Frequent days of week
# ggplot(rossman) +
#     geom_boxplot(aes(x = as.character(Promo), y = Sales))
```

```{r}
# Plot a histogram
# ggplot(rossman) +
#   geom_histogram(aes(x = Customers,
#                      fill = factor(Open))) +
#   facet_grid( ~ Open) + # Separate plots for Open = 1 vs. 0
#   theme(legend.position = "none") # Don't show legend
```

```{r}
# Plot a histogram
# ggplot(rossman) +
#   geom_histogram(aes(x = Customers,
#                      fill = factor(Promo))) +
#   facet_grid( ~ Promo) + # Separate plots for Promo = 1 vs. 0
#   theme(legend.position = "none") # Don't show legend
```

```{r}
# Plot a histogram
# ggplot(rossman) +
#   geom_histogram(aes(x = Customers,
#                      fill = factor(StateHoliday))) +
#   facet_grid( ~ StateHoliday) + # Separate plots for StateHoliday = 1 vs. 0
#   theme(legend.position = "none") # Don't show legend
```

```{r}
# Plot a histogram
# ggplot(rossman) +
#   geom_histogram(aes(x = Customers,
#                      fill = factor(SchoolHoliday))) +
#   facet_grid( ~ SchoolHoliday) + # Separate plots for SchoolHoliday = 1 vs. 0
#   theme(legend.position = "none") # Don't show legend
```





# Data cleaning 
```{r}
#Separate Date into Year Month Day
# 
# rossman_date <- cbind(rossman, year=year(rossman$Date),
#                       month=month(rossman$Date),
#                       day=day(rossman$Date))

```




# Exploratory data analysis

**Dates**

```{r sales-time-series}

# ggplot(full_store, aes(x = Date, y = Sales)) + 
#   geom_area(aes(color = Sales,
#                 fill = Sales), 
#             alpha = 0.4,
#             position = position_dodge(0.7)) +
#   scale_color_manual(values = c("#00AFBB")) +
#   scale_fill_manual(values = c("#00AFBB"))+
#   stat_smooth(
#   color = "#FC4E07", fill = "#FC4E07",
#   method = "loess")
#   
```


**Factors**

```{r sales-per-day-of-week}
# Plot a histogram for DayOfWeek
ggplot(full_store) +
  geom_histogram(aes(x = Sales,
                     fill = DayOfWeek)) +
  facet_wrap(~ DayOfWeek,  ncol = 2) + # Separate plots for DayOfWeek
  theme_minimal()+
  theme(legend.position = "none")+
  labs(title="Sales per day of the week",
       x="Sales",
       y="Count") # Don't show legend
```

```{r sales-on-open-days}
# Plot a histogram for Open
ggplot(full_store) +
  geom_histogram(aes(x = Sales,
                     fill = Open)) +
  facet_wrap(~ Open,  ncol = 2) + # Separate plots for Open
  theme_minimal()+
  theme(legend.position = "none")+
  labs(title="Sales on open days",
       x="Sales",
       y="Count") # Don't show legend
```

```{r sales-on-promo-days}
# Plot a histogram for Promo
ggplot(full_store) +
  geom_histogram(aes(x = Sales,
                     fill = Promo)) +
  facet_wrap(~ Promo,  ncol = 2) + # Separate plots for Promo
  theme_minimal()+
  theme(legend.position = "none")+
  labs(title="Sales per promotion day",
       x="Sales",
       y="Count") # Don't show legend
```

```{r sales-on-school-holidays}
# Plot a histogram for SchoolHoliday
ggplot(full_store) +
  geom_histogram(aes(x = Sales,
                     fill = SchoolHoliday)) +
  facet_wrap(~ SchoolHoliday,  ncol = 2) + # Separate plots for SchoolHoliday
  theme_minimal()+
  theme(legend.position = "none")+
  labs(title="Sales on school holidays",
       x="Sales",
       y="Count") # Don't show legend
```

```{r sales-per-store-type}
# Plot a histogram for StoreType
ggplot(full_store) +
  geom_histogram(aes(x = Sales,
                     fill = StoreType)) +
  facet_wrap(~ StoreType,  ncol = 2) + # Separate plots for StoreType
  theme_minimal()+
  theme(legend.position = "none")+
  labs(title="Sales per store type",
       x="Sales",
       y="Count") # Don't show legend
```

```{r sales-per-assortment-type}
# Plot a histogram for Assortment
ggplot(full_store) +
  geom_histogram(aes(x = Sales,
                     fill = Assortment)) +
  facet_wrap(~ Assortment,  ncol = 3) + # Separate plots for Assortment
  theme_minimal()+
  theme(legend.position = "none")+
  labs(title="Sales per assortment type",
       x="Sales",
       y="Count") # Don't show legend
```

```{r sales-per-ongoing-promotion}
# Plot a histogram for Promo2
ggplot(full_store) +
  geom_histogram(aes(x = Sales,
                     fill = Promo2)) +
  facet_wrap(~ Promo2,  ncol = 2) + # Separate plots for Promo2
  theme_minimal()+
  theme(legend.position = "none")+
  labs(title="Sales per ongoing promotion",
       x="Sales",
       y="Count") # Don't show legend
```

**Numerical variables**

```{r sales-vs-customers}
ggplot(full_store)+
  geom_point(aes(x=Customers,
                 y=Sales),
             alpha=0.3)+
  theme_minimal()+
  labs(title="Sales vs. Customers",
       x="Customers",
       y="Sales") # Don't show legend
```

```{r sales-vs-competitor-distance}
ggplot(full_store)+
  geom_point(aes(x=CompetitionDistance,
                 y=Sales),
             alpha=0.3)+
  theme_minimal()+
  labs(title="Sales vs. competitor distance",
       x="Competitor distance",
       y="Sales") # Don't show legend
```




# Techniques used



# Questions from this analysis

```{r}
# #Calculating annual sales and customers per per store
# annual_sales <-  aggregate (Sales ~ year + Store,
#                             data = rossman_date,
#                             sum)%>%
#   spread(key=year,
#          value = Sales)
# 
# #Changing names to R friendly names
# 
# colnames(annual_sales) <- c("Store",
#                             "sales_2013",
#                             "sales_2014",
#                             "sales_2015")
```

```{r}
# #Joining store characteristics with sales data
# full_store <- left_join(annual_sales,store,by="Store")
```

```{r}
# #Checking linear relationship
# ggplot(annual_sales, aes(x=sales_2013,
#                          y=sales_2014))+geom_point()
# 
```

```{r}
# #Predicting sales one year based on the previous years
# salesSimpleModel <- lm(sales_2014 ~ sales_2013, 
#                        data = annual_sales)
```

```{r}
# #Looking at the distribution of each variable
# ggplot(full_store,aes(x=CompetitionDistance,y=sales_2015))+
#   geom_point()
```
